---
title: "Zambia nightlights data analysis"
output: html_notebook
---

We first need to bring the results of the GEE analysis to the Data folder of this project.
```{r, engine = 'bash', echo = F, eval = FALSE}
#7z l ~/Downloads/Data-20190719T081856Z-001.zip
#7z x -aoa	~/Downloads/Data-20190719T081856Z-001.zip Data/

mv ~/Downloads/hansen_buffered.geojson Data/
mv ~/Downloads/night_buffered.geojson Data/

```


```{r packages, message=FALSE}
rm(list = ls())
gc()
library(tidyverse)
```



### Creating forest loss table

This code creates the table of pixel count per year of forest loss. In other words, it has the number of pixels lost in each year in each cluster.

I used the cluster center coordinates recorded in 2015 (not 2012).

``` {r forest lost}
#### TRY WITH jsonlite PACKAGE ####
forestloss <- jsonlite::fromJSON('Data/hansen_buffered.geojson') 
forestloss <- forestloss$features$properties %>%
  as_tibble() %>% 
  mutate(year = as.integer(imageId) + 2012,
         arealoss_ha = sum / 10000) %>% 
  select(cluster, year, arealoss_ha)

head(forestloss)
```
``` {r}
g <- ggplot(data = forestloss, aes(year, arealoss_ha, group=cluster)) +
  geom_line(alpha=0.1)
g
```


### Creating nightlights time-series table

The nighlight dataset is very variable. Probably due to cloud coverage or errors with the products. Here are the plots of nightlight values sum per cluster|year to see whic month is the most reasonable (less variable, more consistent)/

```{r plot all nightlight years, echo = FALSE}
for (y in seq(1,12,1)){
  #y <- 1
  #### TRY WITH jsonlite PACKAGE ####
  triplets <- jsonlite::fromJSON(paste0('Data/night_buffered', y, '.geojson') )
  #9 and 10 do not go to zero. They are more balanced
  
  night <- triplets$features$properties %>% 
    as_tibble() %>% 
    mutate(year = substr(imageId, start=1, stop = 4) %>% as.integer(),
           urban_degree = sum) %>% 
    arrange(cluster, year)
  
  g <- ggplot(data=night, aes(y = log(urban_degree), x=year, group=factor(cluster))) +
    geom_line(alpha = 0.1) +
    ggtitle(paste0('Year = ', y, '. Urban degree per year per cluster'))
  print(g)
}

```


I chose October. The column of sum is the sum of the radiance values of the buffered area. Could
be considered as a degree of urbanization of the buffered area. For clarity, I rename this column to be
*nightlight_sum*.

``` {r nightlights}
triplets <- jsonlite::fromJSON('Data/night_buffered10.geojson') 

night <- triplets$features$properties %>% 
  as_tibble() %>% 
  mutate(year = substr(imageId, start=1, stop = 4) %>% as.integer(),
         nightlight_sum = sum) %>% 
  arrange(cluster, year)

#estimating urban growth of the period since last year.
night$urban_growth <- c(NA, (tail(night$nightlight_sum, -1) - head(night$nightlight_sum, -1)))

#removing data for 2012
night$urban_growth[night$year==2012] <- NA
head(night)

```


Then we join the two tables and select the columns to be used.

``` {r join}
clusters <- full_join(forestloss, night, by = c('cluster', 'year')) %>%
  select(cluster, year, arealoss_ha, nightlight_sum, urban_growth)
head(clusters)
```
# Exploratory analysis

```{r, echo = FALSE}
g <- ggplot(data=clusters, aes(x = log(nightlight_sum), y=log(arealoss_ha))) +
geom_point(alpha = 0.2) +
ggtitle("urban degree vs. area_loss. Each point represents a cluster|year")
g
```


```{r, echo = FALSE}

g <- ggplot(data=clusters, aes(x = log(urban_growth), y=log(arealoss_ha))) +
geom_point(alpha = 0.2) +
ggtitle("urban growth vs forested area lost. Each point represents a cluster-year")
g
```

Here aI try to correlate area loss from previous year to urban growth with any lag. Forest loss ~ to urban growth from x last year.
``` {r lag}
clusters_lag <- clusters %>% arrange(cluster, year)

lag <- 5

clusters_lag$urbangrowth_lag <- c(rep(NA, lag), head(clusters_lag$urban_growth, -lag) )  
clusters_lag$urbangrowth_lag[clusters_lag$year %in% 2012:(2012+(lag-1))] <- NA  #delete values that should have no growth information for that lag
clusters_lag <- clusters_lag %>% filter(!is.na(lag))

head(clusters_lag)

g <- ggplot(data=clusters_lag, aes(x = log(urbangrowth_lag), y=log(arealoss_ha))) +
  geom_point(alpha = 0.2) +
  ggtitle(paste0('urban growth (lag of ',  lag,' years) vs forested area lost. Each point represents a cluster-year'))
g
```
